\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-11-22 paper/main.tex
ensure fig 4 comes before partial snippets
minor wording changes

Across different versions of a given package, particular modules, functions, and other objects 
may be updated, removed, renamed, or otherwise altered in some manner.  In addition to changing the
behaviors of active computations, these changes can render saved objects created using one
version of a package incompatible with other versions of the
same package.  For example, the popular \texttt{pandas}~\cite{McKi10}
library used to include the \texttt{Panel} data structure for storing 3-dimensional arrays.  
Since version 0.20.0, however, the \texttt{Panel} class has been deprecated, and in 
version 0.25.0, it was removed entirely. Suppose a user had a dataset stored in a \texttt{Panel} object
(created using an older version of \texttt{pandas}) and had saved it to their disk (e.g., for later reuse
or to share with other users) by serializing the \texttt{Panel} with Python's \texttt{pickle} protocol.
The \texttt{pickle} protocol is a popular choice for persisting data in Python, allowing users to save, share, and load arbitrary objects, though with an important caveat:
In order to successfully ``unpickle'' (i.e., load and restore) a ``pickled'' object, its class must be defined in and importable from the same module as when it was saved.
Thus, because of the \texttt{Panel} class's removal, the user's dataset could not be read by any version of \texttt{pandas} from 0.25.0 or beyond. These
incompatibilities are also not limited solely to traditional forms of data.
For example, saved model states and other objects may reference
modules, functions, attributes, classes, or other objects that may not be identical (or even present)
across all versions of their associated package.

@@

@@

After installing and importing \texttt{davos} (lines 1--2), we first smuggle two
utilities for interacting with local files in the code below. The
\texttt{smuggle} statement in line 4 loads the \texttt{is\_file()}
function from the Python standard library's \texttt{os.path}
module. Standard library modules are included with all Python
distributions, so this line is functionally equivalent to an
\texttt{import} statement and does not need or benefit from an onion
comment. Line 5 loads the \texttt{joblib} package~\cite{Varo10},
installing it first, if necessary. Since \texttt{joblib}'s I/O
interface has historically remained stable and backwards-compatible
across releases, requiring that users have a particular exact version
installed would likely be unnecessarily restrictive. However, a
\textit{future} release might introduce some breaking change.  The
onion comment in line 5 helps ensure the analysis notebook continues
to run properly in the future by limiting allowable versions to those
already released when the code was written:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example1}
\end{center}
Line 7 then uses the \texttt{davos.config} object to enable
\texttt{davos}'s \texttt{auto\_rerun} option before smuggling the next
two packages: \texttt{NumPy}~\cite{HarrEtal20} and
\texttt{pandas}. Because these packages rely heavily on custom C data
types, loading the particular versions from the onion comments may
require restarting the notebook kernel if different versions had been previously
imported during the same interpreter session (see
Sec.~\ref{subsec:config}).
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example2}
\end{center}
Setting the \texttt{auto\_rerun} attribute to \texttt{True} is particularly useful
for managing the installation of \texttt{pandas} in the next
lines:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example3}
\end{center}
If we suppose that the data contained in \texttt{data-old.pkl} is
stored in a \texttt{Panel} object, then we must use a version of 
\texttt{pandas} prior to 0.25.0 (i.e., the version in which the \texttt{Panel} 
class was removed) to be able to load it in. Line 11 ensures
that an older version of \texttt{pandas} will be imported, enabling
the data to be read in (and, in line 13, written to a CSV
file, which is compatible with newer versions of \texttt{pandas}).

>>>>

@@

Across different versions of a given package, particular modules, functions, and other objects
may be updated, removed, renamed, or otherwise altered in some manner.  In addition to changing the
behaviors of active computations, these changes can render saved objects created using one
version of a package incompatible with other versions of the
same package.  For example, the popular \texttt{pandas}~\cite{McKi10}
library used to include the \texttt{Panel} data structure for storing 3-dimensional arrays.
Since version 0.20.0, however, the \texttt{Panel} class has been deprecated, and in
version 0.25.0, it was removed entirely. Suppose a user had a dataset stored in a \texttt{Panel} object
(created using an older version of \texttt{pandas}) and had saved it to their disk (e.g., for later reuse
or to share with other users) by serializing the \texttt{Panel} with Python's \texttt{pickle} protocol.
The \texttt{pickle} protocol is a popular, built-in method of persisting data in Python, allowing users to save, share, and load arbitrary objects.
However, in order to successfully ``unpickle'' (i.e., load and restore) a ``pickled'' object, the object's class must be defined in and importable from the same module as when it was saved.
Thus, because of the \texttt{Panel} class's removal, the user's dataset could not be read by any version of \texttt{pandas} from 0.25.0 or beyond. These
incompatibilities are also not limited solely to traditional forms of data.
For example, saved model states and other objects may reference
modules, functions, attributes, classes, or other objects that may not be identical (or even present)
across all versions of their associated package.

@@

After installing and importing \texttt{davos} (lines 1--2), we first smuggle two
utilities for interacting with local files in the code below. The
\texttt{smuggle} statement in line 4 loads the \texttt{is\_file()}
function from the Python standard library's \texttt{os.path}
module. Standard library modules are included with all Python
distributions, so this line is functionally equivalent to an
\texttt{import} statement and does not need or benefit from an onion
comment. Line 5 loads the \texttt{joblib} package~\cite{Varo10},
installing it first, if necessary. Since \texttt{joblib}'s I/O
interface has historically remained stable and backwards-compatible
across releases, requiring that users have a particular exact version
installed would likely be unnecessarily restrictive. However, a
\textit{future} release might introduce some breaking change.  The
onion comment in line 5 helps ensure the analysis notebook continues
to run properly in the future by limiting allowable versions to those
already released when the code was written:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example1}
\end{center}
Line 7 then uses the \texttt{davos.config} object to enable
\texttt{davos}'s \texttt{auto\_rerun} option before smuggling the next
two packages: \texttt{NumPy}~\cite{HarrEtal20} and
\texttt{pandas}. Because these packages rely heavily on custom C data
types, loading the particular versions from the onion comments may
require restarting the notebook kernel if different versions had been previously
imported during the same interpreter session (see
Sec.~\ref{subsec:config}).
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example2}
\end{center}
Setting the \texttt{auto\_rerun} attribute to \texttt{True} is particularly useful
for managing the installation of \texttt{pandas} in the next
lines:
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example3}
\end{center}
If we suppose that the data contained in \texttt{data-old.pkl} is
stored in a pickled \texttt{Panel} object, then we must use a version of
\texttt{pandas} prior to 0.25.0 (i.e., the version in which the \texttt{Panel}
class was removed) to be able to load it in. Line 11 ensures
that an older version of \texttt{pandas} will be imported, enabling
the data to be read in (and, in line 13, written to a CSV
file, which is compatible with newer \texttt{pandas} versions).
